---
title: "CRS and projections"
output: html_document
---

```{r setup, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
library(eurostat)
library(giscoR)
library(sf)
library(spData)
library(RCzechia)
library(terra)
```


--- 

# Goal of this script

* CRS  
* WKT  
* Reprojecting maps (`sf`)
* Reprojecting rasters (`terra`)

---


# Reprojecting maps (`sf`)




## Retrieving spatial information

```{r}
# Load dataset
data("alaska", package = "spData")

#------------------------------------------------------------------------------
# Basic spatial information
#------------------------------------------------------------------------------

alaska

st_geometry_type(alaska)
st_crs(alaska)
st_bbox(alaska)
nrow(alaska)

```

## WKT

```{r}

#------------------------------------------------------------------------------
# WKT representation
#------------------------------------------------------------------------------

wkt <- st_as_text(st_geometry(alaska)[1])
substr(wkt,1,300) # first 300 characters

```

## Reprojecting vector-based maps

```{r}
# WGS84 geographic coordinates
alaska_4326 <- st_transform(alaska, 4326)

# Native projection
alaska_3338 <- alaska

# Lambert Azimuthal Equal Area
alaska_laea <- st_transform(
  alaska,
  "+proj=laea +lat_0=60 +lon_0=-150 +datum=WGS84 +units=m +no_defs"
)

alaska_4326
alaska_3338
alaska_laea

```


## Alaska plotted in different projections

```{r}
ggplot(alaska_4326) +
  geom_sf(fill = "lightgreen", color = "black") +
  ggtitle("WGS84 (EPSG:4326)") +
  theme_minimal()

ggplot(alaska_3338) +
  geom_sf(fill = "lightblue", color = "black") +
  ggtitle("NAD83(NSRS2007)\nAlaska Albers") +
  theme_minimal()

ggplot(alaska_laea) +
  geom_sf(fill = "salmon", color = "black") +
  ggtitle("Lambert Azimuthal\nEqual Area") +
  theme_minimal()

```


## Supervised exercise: Sweden 

* task setup

```{r}
# task setup: download NUTS2 regions for Sweden
# - 
nuts2 <- gisco_get_nuts(
  year = "2024",
  nuts_level = 2,
  epsg = 4326, # or  3035 <--> ETRS89 / LAEA Europe
  resolution = "03"
)
nuts2

Sweden <- nuts2 |> 
  filter(CNTR_CODE== "SE")
Sweden

Sweden |> 
  mutate(NUTS_ID = as.factor(NUTS_ID)) |> 
  ggplot() +
    geom_sf(aes(fill = NUTS_ID), color = "black") +
    theme_minimal()

# How to set the centerpoint for the LAEA projection
ctr <- st_coordinates(st_centroid(Sweden)) 
ctr

# Use geographical center of Sweden
ctr <- Sweden |> 
  st_union() |> # join all NUTS2 regions >> NUTS0
  st_centroid() |> 
  st_coordinates()
ctr

# To use bounding box can be convenient for some countries/geometries...
# (what happens if we use bbox-based center to LAEA project Alaska?)

# What is a bounding box?
bbox_ctr <- Sweden |> 
  st_bbox() |> # generate a vector of 4 elements: xmin, ymin, xmax, ymax
  st_as_sfc()

plot(bbox_ctr, reset=FALSE)
plot(Sweden[,"NUTS_ID"], add=TRUE)

# Find bbox-based center for Sweden
bbox_ctr <- Sweden |> 
  st_bbox() |> # generate a vector of 4 elements: xmin, ymin, xmax, ymax
  st_as_sfc() |> # convert bbox to polygon
  st_centroid() |> # find geographic center
  st_coordinates() # return coordinates
bbox_ctr
```

* Reproject Sweden to LAEA

```{}


```

--- 

# Reprojecting rasters (`terra`) 

* same principle as vector-based projections / reprojections  
* for `terra`, use `project()` instead of `st_transform()`

```{r}
CZelevation <- RCzechia::vyskopis(format = "actual")
CZelevation

CZelevation_laea <- project(CZelevation,  "EPSG:3035") # LAEA Europe
CZelevation_laea

CZelevation_Krovak <- project(CZelevation,  "EPSG:5514") # Krovak projection
CZelevation_Krovak

plot(CZelevation)
plot(CZelevation_laea)
plot(CZelevation_Krovak)
```

----  

## Supervised exercise

* Project `CZelevation` to LAEA, but use Iceland's LAEA reference point: approx 65°N, 19°W (Note: coordinates on the western hemishere should be entered as negative numbers: `lon_0 = -19`)


```{r}
# Complete the task here
```


