---
title: "Map-based operations"
output: html_document
---

```{r setup, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
library(RColorBrewer)
library(eurostat)
library(giscoR)
library(sf)
library(spData)
library(RCzechia)
library(terra)
```

# Goal of this script

> Processing different `sf` geometries  
> Visualizing data: maps, choropleths  
> Extracting information from different types of map data    


--- 

<br>
<br>

## Example 1: Count the number of points within polygons 

* Use North Carolina map  
* Simulate occurrences of some rare disease over a map  
* Count and visualize the occurrences  

```{r}
nc_map <- st_read(system.file("shape/nc.shp", package="sf")) |> 
  dplyr::select(CNTY_ID,NAME)
nc_map
# Map of North Carolina
ggplot() + geom_sf(data = nc_map)
# Simulate rare disease occurrences across NC
#  type = "random", by_polygon = FALSE: default settings
set.seed(123456)
occurrences <- st_sample(nc_map, size=70, type = "random", by_polygon = FALSE)
# Plot the occurences
ggplot() + geom_sf(data = nc_map) + geom_sf(data = occurrences, col ="red")
# Intersection (first argument map, then points)
intersections <- st_intersects(nc_map, occurrences)
# this datasets maps to nc_maps easily
intersections
# reverse order of arguments
intersections2 <- st_intersects(occurrences, nc_map)
# this one doesn't map to nc_maps..
intersections2
# add occurrence counts to nc_maps
nc_map$occurrences <- lengths(intersections)
nc_map
# plot the result (number of points within polygons)
ggplot(nc_map) +
  geom_sf(aes(fill = occurrences)) +
  scale_fill_gradientn('Occurences \nof disease', colours=brewer.pal(9, "Reds"))
```

--- 

<br>
<br>

## Example 2 (supervised work): Land area and population density in Africa

* Land area in Africa (simplified, based on `spData::world()`)  
* Population counts  
* Population density  
* Country-based & region based analysis  

```{r}
world_data <- st_read(system.file("shapes/world.gpkg", package="spData"))
world_data |> filter(continent == "Africa") |>  dplyr::select(subregion) |> plot()
Africa <- world_data |> filter(continent == "Africa")
Africa
# Country-level population density plot
Africa |> 
  mutate(density=pop/area_km2) |> 
  ggplot() +
  geom_sf(aes(fill = density)) +
  scale_fill_gradientn('Population \ndensity \ncountry \naggregate', colours=brewer.pal(9, "Oranges"))+
  theme_minimal() +
  ggtitle("Population density (country-level aggregates)")

# Task 1: Calculate the total population, area, and population density in Africa


# Task 2: Calculate and visualize population density for different Subregions in Africa
# - aggregate Subregion population and area first, 
# then calculate population density over the Subregions


```


---

<br>
<br>

## Example 3 Coffee production in Africa

* Join data from two datasets (spData::world(), spData::coffee_data())


```{r}
Africa
coffee <- spData::coffee_data
coffee # coffee is table, with no spatial information (besides country name)
# join the two datasets 
AfricaCoffee <- Africa |> 
  left_join(coffee, by = "name_long")
#
# Note - a common problem in map handling
#
st_drop_geometry(Africa[3,])
coffee[7,]
# "Democratic Republic of the Congo" is abbreviated to "Congo, Dem. Rep. of" in coffee dataframe
# clearly, this causes left_join() to miss the DRC as coffee producer.
#
# There are different options to fix this problem
#
# a) directly fix (rewrite) DRC name in one of the datasets to make them match (one-off solution)
# b) use regex-based fix to address multiple occurrences (not present in our example, but common issue)
drc <- stringr::str_subset(world$name_long, "Dem*.+Congo")
drc
# fix all occurrences (and even different types of abbreviations)
coffee$name_long[grepl("Congo,", coffee$name_long)] <- drc
# Join the fixed datasets
AfricaCoffee <- Africa |> 
  left_join(coffee, by = "name_long") |> 
  mutate(across(c(coffee_production_2016,coffee_production_2017), \(x) tidyr::replace_na(x, 0)))
AfricaCoffee
# plot 2017 coffee production
AfricaCoffee |> 
  ggplot()+
  geom_sf(aes(fill=coffee_production_2017),col="grey50") +
  scale_fill_gradientn('Coffee \nproduction \n2017', colours=brewer.pal(9, "Blues"))+
  theme_minimal()
```


* As dataframes and `sf` objects are joined, ordering matters:

```{r}
Africa |> 
  left_join(coffee, by = "name_long") |> 
  class()

coffee |> 
  inner_join(Africa, by = "name_long") |> 
  class()
```


--- 

<br>
<br>

## Example 4 (supervised work): River lengths across regions

* Labe and Vltava (2 major Czech rivers)  
* Total lengths  
* Lengths in different NUTS3 regions (for each NUTS3 region, show river length for Labe & Vltava)

```{r}
rivers <- RCzechia::reky(resolution = "low") |> 
  dplyr::filter(NAZEV %in% c("Labe","Vltava")) |> 
  group_by(NAZEV) |> 
  summarize() |> # removes the navigable/non-navigable distinction
  ungroup()
rivers
NUTS3 <- RCzechia::kraje(resolution = "low")
NUTS3
# plot the rivers
ggplot() + 
  geom_sf(data = NUTS3) + 
  geom_sf(data = rivers, aes(col=NAZEV),lwd=1)+
  theme_minimal()
#
# task 1: calculate total length of each river (Labe, Vltava)
# task 2: calculate the length of each river across NUTS3 regions
# Note and remember:
# EPSG:4326 >> geodetic length on Earth surface (in metres)
# EPSG:5514 >> planar Euclidean length in projected metres



```


