---
title: "German minority on the territory of Czechia"
subtitle: "Spatial interpolation using 1930 & 2011 census data"
author: ""
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, message=FALSE, warning=FALSE, echo=FALSE}
library(sf)
library(dplyr)
library(tidyr)
library(RColorBrewer)
library(ggplot2)
library(gridExtra)
library(tibble)
library(knitr)
```

> Goal of the script:  
Evaluate the compound effects of ethnic cleansing of the German population after World War 2, and other post-war influences, on the relative presence on the German minority on the territory of Czechia.

* Analyze changes in relative presence of the German minority.  
* Compare 2011 data (77 LAU1 regions) with 1930 data (330 judicial districts)  

--- 

* We have two spatial datasets (`sf`) with different administrative geographies. Both datasets contain demographic data:  
    - `okresy_1930`: with 330 historical juditcial districts and 1930 demographic data.  
    - `okresy_2011`: with 77 LAU1 districts and 2011 demographic data.  
* 2011-based declaration of ethnicity is based on self-proclamation   
* For information on historical data, refer to:  
    - [https://cuni.maps.arcgis.com/sharing/rest/content/items/c2f19cd1146747a9a8daf5b900e7747b/data](https://cuni.maps.arcgis.com/sharing/rest/content/items/c2f19cd1146747a9a8daf5b900e7747b/data)    
    - [https://karolinum.cz/casopis/auc-geographica/rocnik-50/cislo-2/rok-2015/clanek-2194](https://karolinum.cz/casopis/auc-geographica/rocnik-50/cislo-2/rok-2015/clanek-2194)  
    - Data available through `RCzechia::historie()`  


```{r, include=FALSE}
#
# you can read the data from RCzechia:
#
# sf_use_s2(use_s2 = TRUE) 
# okresy_1930 <- historie("okresy_1930")
# #  If you run st_simplify, the input data is specified with long-lat coordinates 
# # and sf_use_s2() returns TRUE, then the value of dTolerance must be specified in meters.
# simple <- st_simplify(okresy_1930,dTolerance=100)
# simple <- simple |> 
#   dplyr::select("počet obyvatel přítomných","národnost německá","OBJECTID","NAZEV") |> 
#   dplyr::rename("pTot1930" = "počet obyvatel přítomných",
#                 "pGer1930" = "národnost německá",
#                 "id1930" = "OBJECTID",
#                 "name" = "NAZEV")
# st_write(simple,"districts_1930.fgb")
# #
# okresy_2011 <- historie("okresy_2011")
# simple11 <- st_simplify(okresy_2011,dTolerance=100)
# simple11 <- simple11 |> 
#   dplyr::select("počet obyvatel obvykle bydlících","národnost německá","OBJECTID","NAZEV") |> 
#   dplyr::rename("pTot2011" = "počet obyvatel obvykle bydlících",
#                 "pGer2011" = "národnost německá",
#                 "id2011" = "OBJECTID",
#                 "name" = "NAZEV")
# st_write(simple11,"districts_2011.fgb")
#
#
# or just read-in the two sf datasets
CZ1930 <- st_read("data/districts_1930.fgb")
CZ2011 <- st_read("data/districts_2011.fgb")
```

--- 

--- 

The variable of interest is the German population share:

\[
S_i = \frac{G_i}{P_i}
\]

where:

- \(G_i\) is German population in district \(i\),
- \(P_i\) is total population in district \(i\),
- \(S_i\) is the German population share.

--- 

--- 

# Basic data wrangling

```{r,echo=TRUE}
# Data preparation: district area, population density, German shares per region
CZ1930 <- CZ1930 |> 
  mutate(sGer1930 = (pGer1930/pTot1930)*100) |>  # in %
  arrange(id1930) |> 
  st_make_valid() |>
  st_transform(5514) |> # planar projection for area calculation
  mutate(area1930 = as.numeric(st_area(geometry))) |> 
  mutate(area1930 = area1930*1e-06) |> # m2 >> km2
  mutate(pTdens1930 = pTot1930/area1930) |> # calculate total population density
  st_transform(4326) # back to WGS 84
CZ1930

CZ2011 <- CZ2011 |> 
  mutate(sGer2011 = (pGer2011/pTot2011)*100) |>  # in %
  arrange(id2011)|> 
  st_make_valid() |>
  st_transform(5514) |> 
  mutate(area2011 = as.numeric(st_area(geometry))) |> 
  mutate(area2011 = area2011*1e-06) |>
  mutate(pTdens2011 = pTot2011/area2011) |> # calculate total population density
  st_transform(4326)
CZ2011
# Prepare scales that are fixed and can be used across maps, prepare summaries and plots
#
# Population share in 1930, 0-100 %
my_scale <- scale_fill_gradient(
  low = "white",
  high = "red",
  limits = c(0, 100),
  breaks = c(0, 25, 50, 75, 100),
  labels = scales::percent_format(scale = 1, accuracy = 1),
  name = "Germans\n(ethnic)\nshare"
)

# Population share in 2011, 0-3 %
my_scale2 <- scale_fill_gradient(
  low = "white",
  high = "blue",
  limits = c(0, 3),
  breaks = c(0, 1, 2, 3),
  labels = scales::percent_format(scale = 1, accuracy = 1),
  name = "Germans\n(ethnic)\nshare"
)

# Population density (for compatibility across 1930 and 2011)
my_scale3 <- scale_fill_gradientn(
  colours = c("white", "#c7e9c0", "#74c476", "#238b45", "#00441b"),
  trans = scales::pseudo_log_trans(base = 10),
  limits = c(0, 8000),  
  breaks = c(0, 10, 100, 1000, 8000),
  labels = scales::label_number(accuracy = 1),
  name = "Population \nper km sq.")


G1930 <- CZ1930 |> 
  ggplot() +
  geom_sf(aes(fill = sGer1930), color = "grey30", linewidth = 0.1) +
  my_scale +
  labs(title = "1930: The relative share of German population \nscale 0% - 100%") +
  theme_minimal()

G2011 <- CZ2011 |> 
  ggplot() +
  geom_sf(aes(fill = sGer2011), color = "grey30", linewidth = 0.1) +
  my_scale2 +
  labs(title = "2011: The relative share of German population \nscale 0% - 3%") +
  theme_minimal()

Dens1930 <- CZ1930 |> 
  ggplot() +
  geom_sf(aes(fill = pTdens1930), color = "grey30", linewidth = 0.1) +
  my_scale3 +
  labs(title = "1930: Total population density (inhabitants / km2, log-transf.)") +
  theme_minimal()

Dens2011 <- CZ2011 |> 
  ggplot() +
  geom_sf(aes(fill = pTdens2011), color = "grey30", linewidth = 0.1) +
  my_scale3 +
  labs(title = "2011: Total population density (inhabitants / km2, log-transf.)") +
  theme_minimal()

summary_1930 <- CZ1930 |>
  st_drop_geometry() |>
  summarise(
    n_districts = n(),
    Total_population = sum(pTot1930, na.rm = TRUE),
    German_population = sum(pGer1930, na.rm = TRUE),
    German_share_percentage = (German_population / Total_population)*100)

summary_2011 <- CZ2011 |>
  st_drop_geometry() |>
  summarise(
    n_districts = n(),
    Total_population = sum(pTot2011, na.rm = TRUE),
    German_population = sum(pGer2011, na.rm = TRUE),
    German_share_percentage = (German_population / Total_population)*100)
```


---

## Total population, 1930 & 2011 

```{r}
kable(summary_1930, digits = 4, caption = "Summary of 1930 data")
kable(summary_2011, digits = 4, caption = "Summary of 2011 data")
```

**Population density (per km sq.)**

```{r, out.width = "100%", fig.align = "center"}
# Population density (per km square)
summary(CZ1930$pTdens1930)
summary(CZ2011$pTdens2011)
# We use log-transformation to show population density 
# (positive values, large variability for direct color-coding)
grid.arrange(grobs=list(Dens1930,Dens2011),ncol=1)
```

---

## German population, 1930 & 2011

```{r, out.width = "100%", fig.align = "center"}
summary(CZ1930$sGer1930)
summary(CZ2011$sGer2011)
grid.arrange(grobs=list(G1930,G2011),ncol=1)
```

---

* **Discuss aggregated (country-level) population levels and regional distribution for both the total population and German population.**

--- 

# Data & methodology issues

**To evaluate the relative changes in population share of German minority, we face 2 main questions/problems:**

* 1930 and 2011 districts are not compatible.  
* Is it OK to interpolate population shares directly, or do we have to interpolate population levels first?

--- 

## Problem 1

> The 1930 and 2011 districts/polygons are not directly comparable. Therefore, the 1930 values must first be harmonised to the 2011 geography.

To illustrate Problem 1, we can zoom-in to the Western regions of Czechia:

```{r, out.width = "100%", fig.align = "center"}
# note how this single ggplot image draws from two different sf-type dataframes
# CRS compatibility is crucial
CZ2011 |> 
  ggplot() +
  geom_sf(fill = "white", color = "grey30", linewidth = 1) + # 2011 borders
  geom_sf(data=CZ1930, fill = NA, color = "red", linewidth = 0.1 )+ # 1930 borders
  coord_sf(xlim = c(12, 13.5), ylim = c(49.5, 50.5))+ # map range
  theme_minimal()
```

- The mapping of 1930-to-2011 regions is not hierarchical (such as in the NUTS system).  
- A single 1930-type region can intersect with 2 or more 2011 districts.  
- **Area-weighted interpolation** is possible.  
- However, it **assumes that that population is uniformly distributed within each of the 1930 districts**.  
- Actual population (cities, other settlements) is spatially uneven, which affects accuracy of the results.  
- Without higher granularity in data, this cannot be easily solved.  
- We are interpolating from smaller 1930 regions to larger 2011, which alleviates the problem to some extent (the actual bias depends on empirical data properties).  
- Our calculations should be interpreted as approximate harmonised values, rather than exact reconstructions.  

--- 

## Problem 2

> Should we interpolate German population ratios/shares, or should we interpolate the observed populations (total, German) and *then* calculate shares from interpolated population values?

* If the underlying data (`Population_total`, `Population_German`) are available, **always** interpolate underlying data firts, and then calucate shares from the interpolation.  

* In **Problem 1**, we assume uniform distribution of the population **within each 1930-district**. Should we choose to interpolate population shares (ratios) directly, we are also assuming that population density is **uniformly distributed across 1930-districts**.  

---- 

## Ilustrative example 


> Interpolating shares vs. interpolating counts from smaller to higher-level regions


Suppose there are two 1930 districts, `A` and `B`, that together map to the 2011 district `C`.

* Both `A` and `B` contain the same number of minority residents, but district `B` has a much larger total population.  

* The two districts differ in area: district `A` accounts for 40 percent of the target area, while district `B` accounts for 60 percent.  

* District `B` overlaps beyond `C`, but `C` is completely filled by `A` and part of `B`. 

* The rest of `B` (not in `C`) belongs to some district `D` that is not part of our example. 

```{r example-table, echo=FALSE}
example_data <- tibble(
  District = c("A", "B","C"),
  Area = c(40,90,100),
  Area_in_C = c(40,60,100),
  Minority = c(100, 100,NA),
  Population = c(100, 1000,NA),
  Share = Minority / Population
)

kable(
  example_data,
  digits = 3,
  caption = "Mapping of districts A and B to C (and D)"
)
```

```{r, echo=F, out.width = "100%", fig.align = "center"}
library(ggplot2)
library(dplyr)
library(tibble)
# Source districts from 1930
source_districts <- tibble(
  district = c("A", "B"),
  xmin = c(0, 4),
  xmax = c(4, 14),
  ymin = c(0, 0),
  ymax = c(4, 4),
  minority = c(100, 100),
  population = c(100, 1000),
  share = minority / population
)

# Target districts from 2011
target_districts <- tibble(
  district = c("C", "D"),
  xmin = c(0, 10),
  xmax = c(10, 14),
  ymin = c(0, 0),
  ymax = c(4, 4)
)

# Minimal internal labels
weights <- tibble(
  label = c("A: 40 (40% of C)", "B: 60 (60% of C)", "rest of B: 30"),
  x = c(2, 7, 12),
  y = c(2, 2, 2)
)

ggplot() +
  # 1930 source districts
  geom_rect(
    data = source_districts,
    aes(
      xmin = xmin, xmax = xmax,
      ymin = ymin, ymax = ymax,
      fill = district
    ),
    color = "grey30",
    linewidth = 0.5,
    alpha = 0.65
  ) +
  
  # 2011 target districts as outlines
  geom_rect(
    data = target_districts,
    aes(
      xmin = xmin, xmax = xmax,
      ymin = ymin, ymax = ymax
    ),
    fill = NA,
    color = "black",
    linewidth = 1.1,
    linetype = "dashed"
  ) +
  
  # 2011 district labels above plot
  geom_text(
    data = target_districts,
    aes(
      x = (xmin + xmax) / 2,
      y = ymax + 0.45,
      label = district
    ),
    fontface = "bold",
    size = 5
  ) +
  
  # Minimal labels inside rectangles
  geom_text(
    data = weights,
    aes(x = x, y = y, label = label),
    size = 4.5,
    fontface = "bold"
  ) +
  
  # Separator between C and D
  geom_vline(
    xintercept = 10,
    linewidth = 0.8,
    linetype = "dotted"
  ) +
  
  scale_fill_manual(
    values = c(
      "A" = "#fdae61",
      "B" = "#abd9e9"
    )
  ) +
  
  coord_equal(
    xlim = c(-0.5, 14.5),
    ylim = c(-0.4, 5.1),
    expand = FALSE
  ) +
  
  labs(
    title = "Map of A, B, and C regions",
    x = NULL,
    y = NULL
  ) +
  
  theme_void() +
  theme(
    plot.title = element_text(
      face = "bold",
      size = 14,
      hjust = 0.5
    ),
    plot.subtitle = element_text(
      size = 11,
      hjust = 0.5
    ),
    legend.position = "none"
  )
```



$\text{The minority share is: } S_i = \frac{M_i}{P_i},$ where \(M_i\) is the minority population and \(P_i\) is the total population in district \(i\).

--- 

<br>
**Direct interpolation of minority shares from A,B to C**  

If we interpolate the shares directly, we also use direct area-based weights $w_i$, such that $w_A =0.40, w_B = 0.60$, and we obtain:

\[
\widehat{S}^{(dir)}_C
=
w_A \times S_A
+
w_B \times S_B
=
0.4 \times 1.0
+
0.6 \times 0.1
=
0.46.
\]

The target district `C` is estimated to have a minority share of **46 %**.


--- 


<br>
**Interpolating counts first**  
To aggretate and interpolate count variables, we weight each observation in  `A`, `B` by region's overlap with `C`

 \[
w_{iC} =
\frac{a_{iC}}{a_i},
\]

where:

- \(a_{iC}\) is the overlap area between $i$th district and district \(C\),
- \(a_i\) is the total area of 1930 district \(i\).

<br>
Hence, $w_{AC} = 40/40 = 1$ and $w_{BC} = 60/90 = 0.667$.

<br>
Next, we interpolate the minority and total population counts separately.

First, interpolate the minority count:

$$
\widehat{M_C} = \sum_{i=1}^N w_{iC}M_i
= w_{AC} \times M_A + w_{BC} \times M_C
= 1 \times 100 + 0.667 \times 100 = 167.
$$

Then interpolate the total population count:

$$
\widehat{P_C} = \sum_{i=1}^N w_{iC}P_i
=
w_{AC} \times P_A + w_{BC} \times P_B = 
1 \times 100 + 0.667 \times 1000
= 767.
$$

Finally, compute the share:

\[
\widehat{S_C}^{(2step)}
=
\frac{\widehat{M_C}}{\widehat{P_C}}
=
\frac{167}{767}
=
0.2177.
\]

The count-based interpolated minority share is approximately **21.8 %**.


--- 

**Interpretation**

* Direct share interpolation method gives district `A` and district `B` influence according to their area weights only. Same population densities are assumed across 1930-regions.  

* Counts-first method weights account for both area and population sizes (total, minority). Here, we only assume constant population densities within 1930-regions.  

----

# Intersections between 1930 and 2011 regions

Checklist before starting the interpolation (all OK in our sandbox example, not typical for actual empirical data):  

* Polygon (geometry) validity  
* CRS compatibility  
* Use planar projection to calculate polygon areas. 


> In the `intersections` object, each row (intersected polygon) represents the area shared by one 1930-district and one 2011-district.

```{r intersections, warning=FALSE}
# for consistency of weight calculations 
# (1930 and 2011 areas calculated using EPSG 5514)
CZ1930 <- CZ1930 |> st_transform(5514)
CZ2011 <- CZ2011 |> st_transform(5514)

intersections <- st_intersection( # two sf objects follow, we select columns
  CZ1930 |> dplyr::select(id1930,pTot1930,pGer1930,area1930),
  CZ2011 |> dplyr::select(id2011)
) |>
  mutate(
    area_intersection = (as.numeric(st_area(geometry)))*1e-6, 
    # calculate area of intersection based on EPSG 5514
    # convert area to km2 for consistency
    w_area = area_intersection / area1930) # area1930 in km sq.
#
#
# back to WGS 84
CZ1930 <- CZ1930 |> st_transform(4326)
CZ2011 <- CZ2011 |> st_transform(4326)
# show the intersections object
intersections
```

Here,

 \[
\text{w_area  =  }
w_{ij} =
\frac{a_{ij}}{a_i},
\]

where (following from our example above):

- \(a_{ij}\) is the overlap area between 1930 district \(i\) and 2011 district \(j\),
- \(a_i\) is the total area of 1930 district \(i\).

Note that we are going to "translate" 1930 districts to 2011 districts (not the ohter way around).

--- 

# Data interpolation

* German population and total population are **counts**.   
* 1930 data are allocated by area weights to the 2011 data and then summed.  

```{r correct-method}
intersections_sums <- intersections |>
  st_drop_geometry() |> 
  mutate(
    pTot1930_alloc = pTot1930 * w_area,
    pGer1930_alloc = pGer1930 * w_area) |> 
  group_by(id2011) |> # weighted sums: aggregate 1930 data into 2011 regions
  summarise(
    pTot1930_est = sum(pTot1930_alloc, na.rm = TRUE),
    pGer1930_est = sum(pGer1930_alloc, na.rm = TRUE),
    sGer1930_est = (pGer1930_est / pTot1930_est)*100) |> # in %
  ungroup() |> 
  arrange(id2011)
  

Full_df <- CZ2011 |> 
  dplyr::select(id2011,sGer2011,pTot2011,pGer2011) |> 
  left_join(intersections_sums, by = "id2011") |> 
  mutate(percPointChange = (sGer2011 - sGer1930_est))
str(Full_df)
```

--- 

# Map interpolations - Diagnostics

* Check whether population counts are preserved after interpolation 

    - The count-based interpolation (1930 regions - to - 2011 regions) should approximately preserve the original 1930 counts.  
    - Maps are not perfect, there are some rounding errors, minor historical changes in state borders, split of Czechoslovakia, etc.  

```{r diagnostics-total-preservation}
diagnostics_totals <- tibble(
  quantity = c(
    "Original 1930 total population",
    "Interpolated 1930 total population",
    "Original 1930 German population",
    "Interpolated 1930 German population"
  ),
  value = c(
    sum(CZ1930$pTot1930, na.rm = TRUE),
    sum(Full_df$pTot1930_est, na.rm = TRUE),
    sum(CZ1930$pGer1930, na.rm = TRUE),
    sum(Full_df$pGer1930_est , na.rm = TRUE)
  )
)

diagnostics_totals |>
  mutate(value = round(value, 2)) |>
  kable(caption = "Total preservation diagnostics")
```



* Check coverage of 1930 districts in mapping to 2011 regions

    - For each 1930 district, the sum of allocated area weights should be close to 1  
    - Again, maps are not ideal, rounding errors, minor changes in international borders, etc. 

```{r diagnostics-coverage}
coverage_by_1930 <- intersections |>
  st_drop_geometry() |>
  group_by(id1930) |>
  summarise(
    weight_sum = sum(w_area, na.rm = TRUE),
    .groups = "drop"
  )

summary(coverage_by_1930$weight_sum)
```

--- 

# Maps

## German 1930 population share translated to 2011 districts

```{r , fig.width=8, fig.height=6}
g1930Interp <- ggplot(Full_df) +
  geom_sf(aes(fill = sGer1930_est), color = "grey30",linewidth = 0.1) +
  my_scale +
  labs(title = "1930: German population share: Overlay of 1930 data \non 2011 LAU1 districts, scale 0% - 100%") +
  theme_minimal()
grid.arrange(grobs=list(G1930,g1930Interp),ncol=1)
```


## percentage-point change from 1930 to 2011

```{r, fig.width=8, fig.height=6}
ggplot(Full_df) +
  geom_sf(
    aes(fill = percPointChange),
    color = "gray30",
    linewidth = 0.1
  ) +
  scale_fill_gradient2(
    low = "firebrick",
    mid = "white",
    high = "steelblue",
    midpoint = 0,
    name = "Change\np.p."
  ) +
  labs(
    title = "Changes German population share, 1930 to 2011 (percentage point differences)",
    subtitle = "1930 values harmonised to 2011 districts"
  ) +
  theme_minimal()
```

# Alternative representations 

> Total population, German minority, 1930 vs 2011, country-level, regional changes

- The percentage point changes plotted above are valid, yet they may not represent well the true level of decline of the German minority presence in Czechia.  

- We may use the comparison of total population relative changes (1930 to 2011, using 2011 regions) and German population relative changes. 

- Note that the **change in total population (country-level) is very small**, unlike the change in German population:

```{r national-comparison, echo=FALSE}
overall_1930 <- CZ1930 |>
  st_drop_geometry() |>
  summarise(
    year = 1930,
    Tot_population = sum(pTot1930, na.rm = TRUE),
    German_population = sum(pGer1930, na.rm = TRUE),
    share_german = German_population / Tot_population
  )

overall_2011 <- CZ2011 |>
  st_drop_geometry() |>
  summarise(
    year = 2011,
    Tot_population = sum(pTot2011, na.rm = TRUE),
    German_population = sum(pGer2011, na.rm = TRUE),
    share_german = German_population / Tot_population
  )

overall_change <- bind_rows(overall_1930, overall_2011) |>
  mutate(
    share_german_percent = 100 * share_german
  )

overall_change |>
  mutate(across(where(is.numeric), \(x) round(x, 4))) |>
  kable(caption = "National-level German population share")
```

<br>
**Country-level relative changes in population (Total):**

```{r national-change}
# Change in total population in Czechia: 1930 to 2011, in (%):
100 * ((overall_2011$Tot_population - overall_1930$Tot_population)/overall_1930$Tot_population)
```

<br>
**Country-level relative changes in population (German minority):**

```{r}
# Change in German population in Czechia: 1930 to 2011, in (%):
100 * ((overall_2011$German_population - overall_1930$German_population)/overall_1930$German_population)
```

<br>
**Plots of regional relative changes in population (Total & German):**

```{r map-change-correct, fig.width=8, fig.height=6}
Full_df <- Full_df |> 
  mutate(pchTotPop = 100*((pTot2011-pTot1930_est)/pTot1930_est),
         pchGerPop = 100*((pGer2011-pGer1930_est)/pGer1930_est))

my_scale4 <- scale_fill_gradient2(
  low = "firebrick",
    mid = "white",
    high = "steelblue",
    midpoint = 0,
  limits = c(-100, 120),
  breaks = c(-100, -60, 0, 60, 120),
  labels = scales::percent_format(scale = 1, accuracy = 1),
  name = "Relative change \nin population"
)


RCT <- ggplot(Full_df) +
  geom_sf(
    aes(fill = pchTotPop),
    color = "gray30",
    linewidth = 0.1
  ) +
  my_scale4 +
  labs(
    title = "Relative changes in total population, 1930 to 2011 (%)",
    subtitle = "1930 values harmonised to 2011 districts"
  ) +
  theme_minimal()

RCG <- ggplot(Full_df) +
  geom_sf(
    aes(fill = pchGerPop),
    color = "gray30",
    linewidth = 0.1
  ) +
 my_scale4 +
  labs(
    title = "Relative changes in German population, 1930 to 2011 (%)",
    subtitle = "1930 values harmonised to 2011 districts"
  ) +
  theme_minimal()

grid.arrange(grobs=list(RCT,RCG),ncol=1)

# Relative share of German population 1930 vs 2011 (using 2011 regions)
grid.arrange(grobs=list(g1930Interp,G2011),ncol=1)
```


# Assumptions and limitations

The main limiting assumption is **uniform spatial distribution within 1930-source regions**. 

A better approach would use auxiliary information, for example:

- settlement points,
- built-up areas,
- historical cadastral data (land use information),
- elevation or terrain,
- gridded population data.

This would allow for improved weighted interpolation, where population is allocated according to plausible settlement locations rather than uniformly by the 1930-region area.

