---
title: "rasters_presentation"
author: " "
date: " "
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Install spDataLarge from the geocompx R-universe if not available
if (!requireNamespace("spDataLarge", quietly = TRUE)) {
  install.packages(
    "spDataLarge",
    repos = "https://geocompx.r-universe.dev"
  )
}
library(stars)
library(mapview)
library(h3jsr)
library(osmdata)
library(RCzechia)
library(grid)
library(ggplot2)
library(dplyr)
library(tibble)
library(patchwork)
library(viridis)
library(terra)
library(exactextractr)
library(spDataLarge)
```


```{r}
# create a raster object with three layers
r <- rast(ncol=10, nrow=10, xmin=-150, xmax=-80, ymin=20, ymax=60)
set.seed(123456)
values(r) <- rnorm(ncell(r))
r2 <- r * r
r3  <- r > 0
s <- c(r, r2, r3)
names(s)[c(1:3)] <- c("z = rnorm(ncell(r))", "z^2 values","z>0 T/F classification")
s
plot(s,nc=3)
```

```{r}

# ------------------------------------------------------------
# 1. Create a 5x5 raster-like grid
# ------------------------------------------------------------

set.seed(123)

n <- 5

raster_df <- expand.grid(
  col = 1:n,
  row = 1:n
) |>
  as_tibble() |>
  arrange(row, col) |>
  mutate(
    cell_id = row_number(),
    value = c(
      92, 55, 48, 21, 76,
      58, 70, NA, 37, 64,
      NA, 12, 94, 11, 80,
      36, 83,  4, 88, 45,
      20, NA, 67, 73,  9
    )
  )

# Reverse row order for plotting, so row 1 appears at the top
raster_df <- raster_df |>
  mutate(
    y = n - row + 1
  )

# ------------------------------------------------------------
# 2. Common plot theme
# ------------------------------------------------------------

grid_theme <- theme_void() +
  theme(
    plot.title = element_text(
      size = 18,
      hjust = 0,
      face = "plain"
    ),
    plot.margin = margin(10, 15, 10, 15)
  )

# ------------------------------------------------------------
# 3. Plot A: Cell IDs
# ------------------------------------------------------------

p_ids <- ggplot(raster_df, aes(x = col, y = y)) +
  geom_tile(
    fill = "white",
    color = "black",
    linewidth = 0.6
  ) +
  geom_text(
    aes(label = cell_id),
    size = 6
  ) +
  coord_equal() +
  labs(title = "A. Cell IDs") +
  grid_theme

# ------------------------------------------------------------
# 4. Plot B: Cell values
# ------------------------------------------------------------

p_values <- ggplot(raster_df, aes(x = col, y = y)) +
  geom_tile(
    fill = "white",
    color = "black",
    linewidth = 0.6
  ) +
  geom_text(
    aes(label = ifelse(is.na(value), "NA", value)),
    size = 6
  ) +
  coord_equal() +
  labs(title = "B. Cell values") +
  grid_theme

# ------------------------------------------------------------
# 5. Plot C: Colored values
# ------------------------------------------------------------

p_colored <- ggplot(raster_df, aes(x = col, y = y, fill = value)) +
  geom_tile(
    color = NA
  ) +
  scale_fill_gradient2(
    low = "#0b7fb3",
    mid = "#f7f7f7",
    high = "#d4002a",
    midpoint = 50,
    na.value = "grey30",
    guide = "none"
  ) +
  coord_equal() +
  labs(title = "C. Data visualization") +
  grid_theme

# ------------------------------------------------------------
# 6. Combine panels
# ------------------------------------------------------------

p_ids + p_values + p_colored +
  plot_layout(ncol = 3)
```




```{r}
library(terra)
r <- rast(nrows = 5, ncols = 5, xmin = 0, xmax = 5, ymin = 0, ymax = 5)
r
values(r) <- 1:ncell(r)

ncell(r)
values(r)

xyFromCell(r, 1)
xyFromCell(r, 13)
xyFromCell(r, 25)

cellFromXY(r, cbind(x = 2.5, y = 2.5))

# ------------------------------------------------------------
# 1. Create a simple 5x5 raster
# ------------------------------------------------------------

r <- rast(
  nrows = 5,
  ncols = 5,
  xmin = 0, xmax = 5,
  ymin = 0, ymax = 5,
  crs = "local"
)

values(r) <- 1:ncell(r)

# ------------------------------------------------------------
# 2. Create data frame with cell IDs and xy coordinates
# ------------------------------------------------------------

cell_ids <- 1:ncell(r)

cell_df <- as_tibble(xyFromCell(r, cell_ids)) |>
  mutate(
    cell_id = cell_ids,
    row = rowFromCell(r, cell_ids),
    col = colFromCell(r, cell_ids),
    xy_label = paste0("(", x, ", ", y, ")")
  )

cell_df

# ------------------------------------------------------------
# 3. Common theme
# ------------------------------------------------------------

grid_theme <- theme_void() +
  theme(
    plot.title = element_text(
      size = 18,
      hjust = 0.5
    ),
    plot.margin = margin(10, 15, 10, 15)
  )

# ------------------------------------------------------------
# 4. Plot A: Cell IDs
# ------------------------------------------------------------

p_ids <- ggplot(cell_df, aes(x = x, y = y)) +
  geom_tile(
    width = 1,
    height = 1,
    fill = "white",
    color = "black",
    linewidth = 0.6
  ) +
  geom_text(
    aes(label = cell_id),
    size = 6
  ) +
  coord_equal() +
  labs(title = "A. Cell IDs") +
  grid_theme

# ------------------------------------------------------------
# 5. Plot B: Cell-centre xy coordinates
# ------------------------------------------------------------

p_xy <- ggplot(cell_df, aes(x = x, y = y)) +
  geom_tile(
    width = 1,
    height = 1,
    fill = "white",
    color = "black",
    linewidth = 0.6
  ) +
  geom_text(
    aes(label = xy_label),
    size = 2
  ) +
  coord_equal() +
  labs(title = "B. Cell-centre coordinates") +
  grid_theme

# ------------------------------------------------------------
# 6. Combine plots
# ------------------------------------------------------------
p_ids + p_xy
```

---

## Hexagonal tessellation and SpatVector

```{r}
set.seed(123456)
# Example raster
r <- rast(nrows = 100,ncols = 100,
  xmin = 0, xmax = 100,ymin = 0, ymax = 100,
  crs = "local")
r

values(r) <- rnorm(ncell(r),mean=0,sd=2) + 
  seq(from=1,to=5,length.out=ncell(r))

# Create hexagonal polygons over raster extent
hex <- tessellate(r,size = 10,type = "hexagon")
hex
plot(r)
plot(hex, add = TRUE, border = "red",lwd=2)

hex_mean <- extract(r,hex,fun = mean,na.rm = TRUE)
hex_min <- extract(r,hex,fun = min,na.rm = TRUE)
hex$mean_value <- hex_mean[, 2]
hex$min_value <- hex_min[, 2]
hex
plot(hex, "mean_value")

```

--- 

## Subsetting values


```{r}

r <- rast(ncol=10, nrow=10, xmin=-150, xmax=-80, ymin=20, ymax=60)
set.seed(123456)
values(r) <- rnorm(ncell(r))
r2 <- r * r
r3  <- r > 0
s <- c(r, r2, r3)
names(s)[c(1:3)] <- c("z = rnorm(ncell(r))", "z^2 values","z>0 T/F classification")
s
plot(s,nc=3)

# subsetting
r[1]
r[1,1]
s[1,1,1]

s[1,1]
s[1,]
s[]

s[1,1,1] <- NA # sets the first element to "NA"
plot(s,nc=3)

s[1,1] <- NA # compare with the previous command
plot(s,nc=3)

s[1,1] <- c(0.5,0.25,TRUE) 
plot(s,nc=3)
```


```{r}
# spatial subsetting
xyFromCell(r,15)
id <- cellFromXY(r, matrix(c(-118.5,54), ncol =2))
id
r[id]
r[id] <- 1


# For arrays
s[[3]] |> xyFromCell(50)
id <- cellFromXY(s[[3]], matrix(c(-83.5, 42), ncol =2))
id # cell 50 in layer 3
s[[3]][50]

# The above commands retrieve values from cells. For spatial outpus, use:
r_subset <- r[1:20, drop = FALSE]
r_subset
plot(r_subset)

r3x3 <- r # 3x3 subset of r
terra::values(r3x3) <- FALSE
r3x3[c(1:3,11:13,21:23)] <- TRUE
r2_3x3 <- mask(r2,r3x3)
s2 <- c(r,r2,r3x3,r2_3x3)
plot(s2,nc=4)

# compare to NA use
r3x3 <- r # 3x3 subset of r
terra::values(r3x3) <- NA
r3x3[c(1:3,11:13,21:23)] <- TRUE
r2_3x3 <- mask(r2,r3x3)
s2 <- c(r,r2,r3x3,r2_3x3)
plot(s2,nc=4)
s3 <- c(r2,r3x3,r2_3x3)
names(s3)[c(1:3)] <- c("z^2 values","3 x 3 subset","masked values")
plot(s3,nc=3)

# another spatial subset
r3subset <- terra::ifel(r3, TRUE, NA)
r2_spatial_subset <- mask(r2,r3subset)
s4 <- c(r2,r3,r3subset,r2_spatial_subset)
names(s4)[c(1:4)] <- c("z^2 values","z>0 T/F classification", "subset of TRUE","masked values")
plot(s4,nc=4)
# inverse=TRUE switches the masking output
r2_spatial_subset_inv <- mask(r2,r3subset, inverse = TRUE)
s4 <- c(r2,r3,r3subset,r2_spatial_subset_inv)
names(s4)[c(1:4)] <- c("z^2 values","z>0 T/F classification", "subset of TRUE","inverse masked output")
plot(s4,nc=4)
```

--- 

## Combining rasters: mosaic()

```{r}
# Combine adjacent (and maybe partially overlapping) rasters using {terra:mosaic()}
# - say, we download satellite photographs for analysis, yet wee need to combine
#   them into one file for subsequent analysis:
#
# terra help-based example:
# ?terra::mosaic
x <- rast(xmin=-110, xmax=-60, ymin=40, ymax=70, res=1, vals=1)
y <- rast(xmin=-95, xmax=-45, ymax=60, ymin=30, res=1, vals=2)
z <- rast(xmin=-80, xmax=-30, ymax=50, ymin=20, res=1, vals=3)
# x,y,z are 3 different rasters, partly overlapping, 
# "observed" values are also different (1,2,3)
m1 <- mosaic(x, y, z, fun="mean") # fun="mean" is the default setting
m2 <- mosaic(x, y, z, fun="blend") # fun="blend" is the smoother
m3 <- c(m1,m2)
names(m3)[c(1:2)] <- c("mean","blend")
plot(m3,nc = 2,mar = c(2, 2, 2, 4)) 
# note how values in overlapping areas are aggregated

```


--- 


## Local operations

```{r}
# NDVI
multi_raster_file <- system.file("raster/landsat.tif", package = "spDataLarge")
multi_rast <- rast(multi_raster_file)
plot(multi_rast)
multi_rast
multi_rast <- project(multi_rast, "epsg:4326")
names(multi_rast)[c(1:4)] <- c("Blue", "Green","Red","NIR")

# https://www.usgs.gov/faqs/how-do-i-use-a-scale-factor-landsat-level-2-science-products

# https://www.usgs.gov/faqs/what-are-band-designations-landsat-satellites 

multi_rast <- (multi_rast * 0.0000275) - 0.2

summary(values(multi_rast))

# Values now should be in a range between 0 and 1. 
# This is not the case here, probably due to the presence of clouds, snow cover, etc. 
# Fix negative values for RGB pictures (photos)
photoRast <- multi_rast
photoRast[photoRast < 0] = 0

plot(photoRast)

# Color image can be assembled from the RGB layers
plotRGB(
  photoRast,
  r = 3, # layer 3 as red
  g = 2, # layer 2 as green
  b = 1, # layer 1 as blue
  stretch = "lin",
  mar = c(2, 2, 3, 2),
  axes = TRUE,
  main = "Landsat RGB composite"
)

# NDVI calculation using the R and NIR layers
ndviFN <- function(nir, red){
  (nir - red) / (nir + red)
}

ndvi_rast = lapp(multi_rast[[c(4, 3)]], fun = ndviFN)

names(ndvi_rast) <- "NDVI"

# Optional: keep values within theoretical NDVI range
ndvi_rast <- clamp(
  ndvi_rast,
  lower = -1,
  upper = 1,
  values = TRUE
)

# Red-white-green divergent palette
ndvi_pal <- colorRampPalette(
  c("red", "white", "darkgreen")
)(100)

plot(
  ndvi_rast,
  col = ndvi_pal,
  main = "NDVI",
  axes = TRUE
)

#-------------------------------------------------------------------------------
# 
# Prepare NDVI-based classification and plot
#
# NDVI <= 0        Water, clouds, snow, shadows
# 0 < NDVI <= 0.1  Bare soil, built-up land
# 0.1-0.3          Sparse/stressed vegetation
# 0.3-0.5          Moderate vegetation
# NDVI > 0.5       Dense/healthy vegetation

classify_ndvi <- function(ndvi) {
  # Classification matrix:
  # from, to, class_id
  rcl <- matrix(
    c(-Inf, 0.0, 1,
       0.0, 0.1, 2,
       0.1, 0.3, 3,
       0.3, 0.5, 4,
       0.5, Inf, 5),
    ncol = 3,
    byrow = TRUE)
    ndvi_class <- classify(
    ndvi,
    rcl = rcl,
    include.lowest = TRUE,
    right = TRUE)
  ndvi_class <- as.factor(ndvi_class)
  levels(ndvi_class) <- data.frame(
    value = 1:5,
    NDVI_class = c(
      "Water, clouds, snow, shadows",
      "Bare soil, built-up land",
      "Sparse or stressed vegetation",
      "Moderate vegetation",
      "Dense, healthy vegetation"))
  names(ndvi_class) <- "NDVI_class"
  ndvi_class
}

ndvi_cols <- c(
  "Water, clouds, snow, shadows" = "red",
  "Bare soil, built-up land" = "grey80",
  "Sparse or stressed vegetation" = "khaki",
  "Moderate vegetation" = "yellowgreen",
  "Dense, healthy vegetation" = "darkgreen"
)

ndvi_class <- classify_ndvi(ndvi_rast)
ndvi_class

plot(ndvi_class,col = ndvi_cols,main = "NDVI classification")
```

## Focal operations

```{r}

# ------------------------------------------------------------
# 1. Create a small 6 x 6 raster matching the teaching figure
# ------------------------------------------------------------
set.seed(123456)
vals <- matrix(rnorm(36,mean = 0,sd=3),nrow = 6,byrow = TRUE)

r <- rast(
  nrows = 6,
  ncols = 6,
  xmin = 0, xmax = 6,
  ymin = 0, ymax = 6,
  crs = "local"
)

# terra fills cells from upper-left to lower-right
values(r) <- vals # as.vector(t(vals))

# ------------------------------------------------------------
# 2. Apply 3 x 3 focal minimum
# ------------------------------------------------------------

r_focal <- focal(r,w = matrix(1, nrow = 3, ncol = 3),fun = max)

# ------------------------------------------------------------
# 3. Convert rasters to data frames for ggplot
# ------------------------------------------------------------


r_to_df <- function(x, x_offset = 0, panel = "input") {
  
  cell_ids <- seq_len(ncell(x))
  
  nr <- dim(x)[1]
  
  rc <- as.data.frame(rowColFromCell(x, cell_ids))
  names(rc) <- c("row_id", "col_id")
  
  vals <- as.data.frame(values(x))
  names(vals) <- "value"
  
  data.frame(
    cell = cell_ids,
    rc,
    vals
  ) |>
    mutate(
      x_plot = col_id + x_offset,
      y_plot = nr - row_id + 1,
      label = ifelse(is.na(value), "NA", ""),
      panel = panel
    )
}

df_input <- r_to_df(r, x_offset = 0, panel = "input")
df_output <- r_to_df(r_focal, x_offset = 6.6, panel = "output")

df_all <- bind_rows(df_input, df_output)

# Optional: common colour limits for both rasters
value_limits <- range(df_all$value, na.rm = TRUE)

# ------------------------------------------------------------
# 4. Plot the schematic figure
# ------------------------------------------------------------

ggplot(df_all, aes(x = x_plot, y = y_plot)) +
  
  # Raster cells
  geom_tile(
    aes(fill = value),
    color = "grey45",
    linewidth = 0.35,
    width = 1,
    height = 1
  ) +
  
  # Only print NA labels
  geom_text(
    data = df_all |> filter(is.na(value)),
    aes(label = label),
    size = 4,
    color = "black"
  ) +
  
  # Orange 3 x 3 focal window on the input raster
  annotate(
    "rect",
    xmin = 0.5, xmax = 3.5,
    ymin = 3.5, ymax = 6.5,
    color = "orange",
    fill = NA,
    linewidth = 1.4
  ) +
  
  # Orange centre cell on input raster
  annotate(
    "rect",
    xmin = 1.5, xmax = 2.5,
    ymin = 4.5, ymax = 5.5,
    color = "orange",
    fill = NA,
    linewidth = 1.1
  ) +
  
  # Orange output cell receiving the focal value
  annotate(
    "rect",
    xmin = 8.1, xmax = 9.1,
    ymin = 4.5, ymax = 5.5,
    color = "orange",
    fill = NA,
    linewidth = 1.1
  ) +
  
  # Arrow from input window to output value
  annotate(
    "segment",
    x = 3.55, xend = 8.05,
    y = 5.0, yend = 5.0,
    linewidth = 0.7,
    color = "red3",
    arrow = arrow(length = unit(0.25, "cm"))
  ) +
  
  # Viridis colour scale
  scale_fill_viridis_c(
    option = "viridis",
    limits = value_limits,
    na.value = "grey75"
  ) +
  
  coord_equal(
    xlim = c(0.3, 13.0),
    ylim = c(0.3, 6.7),
    expand = FALSE
  ) +
  
  theme_void() +
  theme(
    legend.position = "none",
    plot.margin = margin(10, 10, 10, 10)
  )
```

---

## Zonal operations (Zion park example revisited)

```{r}
z <- terra::zonal(ndvi_rast, ndvi_class, fun = "mean")
class(z)
z
```


--- 

## Global operations 

```{r}
# project for simple distance calc. (m)
ndvi_proj <- project(ndvi_class, "EPSG:32612")

set.seed(12345)
spots <- spatSample(
  ndvi_proj,
  size = 5,
  method = "random",
  as.points = TRUE,
  na.rm = TRUE,
  values = FALSE
)
spots$id <- 1:nrow(spots)
spots

dist_to_spot <- terra::distance(ndvi_proj, spots,method="geo",rasterize=FALSE)
# Convert to kilometres if CRS units are metres
dist_to_spot <- dist_to_spot / 1000
names(dist_to_spot) <- "km distance to nearest spot"

# Original raster layer with random spots
plot(
  ndvi_proj,
  main = "NDVI categorization"
)

points(
  spots,
  pch = 21,
  bg = "red",
  col = "black",
  cex = 1.5
)

text(
  crds(spots),
  labels = spots$id,
  pos = 3,
  cex = 0.9
)

# Distance to nearest spot
plot(
  dist_to_spot,
  main = "Distance to nearest spot (km)"
)

points(
  spots,
  pch = 21,
  bg = "red",
  col = "black",
  cex = 1.5
)

text(
  crds(spots),
  labels = spots$id,
  pos = 3,
  cex = 0.9
)

```


## Extent harmonization

```{r}
set.seed(1234567)

r <- rast(
  nrows = 6,
  ncols = 6,
  xmin = 0, xmax = 6,
  ymin = 0, ymax = 6,
  crs = "local"
)

values(r) <- rnorm(ncell(r))

# 2. Extend raster

# extend(r, c(1, 2)) adds:
# 1 row above and below
# 2 columns left and right
r_ext <- extend(r, c(1, 2))
plot(r_ext) # note the NA values

try(
  r3 <- r + r_ext
)

r_new = extend(r, r_ext)
r_sum <- r_new + r_ext
r_sum

# ------------------------------------------------------------
# 3. Helper function: convert raster to plotting dataframe
# ------------------------------------------------------------

r_to_df <- function(x, x_offset = 0, y_offset = 0, panel = "raster") {
  
  cell_ids <- seq_len(ncell(x))
  
  nr <- dim(x)[1]
  
  rc <- as.data.frame(rowColFromCell(x, cell_ids))
  names(rc) <- c("row_id", "col_id")
  
  vals <- as.data.frame(values(x))
  names(vals) <- "value"
  
  data.frame(
    cell = cell_ids,
    rc,
    vals
  ) |>
    mutate(
      x_plot = col_id + x_offset,
      y_plot = nr - row_id + 1 + y_offset,
      panel = panel
    )
}

# ------------------------------------------------------------
# 4. Convert original and extended rasters to data frames
# ------------------------------------------------------------

gap <- 3

# Center the smaller original raster vertically relative to the extended raster
y_offset_original <- (nrow(r_ext) - nrow(r)) / 2

df_r <- r_to_df(
  r,
  x_offset = 0,
  y_offset = y_offset_original,
  panel = "original"
)

df_ext <- r_to_df(
  r_ext,
  x_offset = ncol(r) + gap,
  y_offset = 0,
  panel = "extended"
)

df_all <- bind_rows(df_r, df_ext)

# Common color scale for both rasters
value_limits <- range(df_all$value, na.rm = TRUE)

# ------------------------------------------------------------
# 5. Plot original and extended raster side by side
# ------------------------------------------------------------

ggplot(df_all, aes(x = x_plot, y = y_plot)) +
  geom_tile(
    aes(fill = value),
    color = "grey25",
    linewidth = 0.55,
    width = 1,
    height = 1
  ) +
  scale_fill_viridis_c(
    limits = value_limits,
    na.value = "grey75"
  ) +
  coord_equal(
    xlim = c(0.5, ncol(r) + gap + ncol(r_ext) + 0.5),
    ylim = c(0.5, nrow(r_ext) + 0.5),
    expand = FALSE
  ) +
  theme_void() +
  theme(
    legend.position = "none",
    plot.margin = margin(20, 20, 20, 20)
  )
```

---

## Aggregation example

```{r}
# Digital elevation model (DEM) of the Mongón study area.
mongon <- rast(system.file("raster/dem.tif", package = "spDataLarge"))
mongon_agg = aggregate(mongon, fact = 6, fun = mean)
# White to blue palette
pal_white_blue <- colorRampPalette(c("white", "darkblue"))(100)
plot(mongon, col = pal_white_blue,main = "Mongón elevation")
plot(mongon_agg, col = pal_white_blue,main = "Mongón elevation aggregated")

```


--- 

## Resampling example

```{r}
r <- rast(nrows=3, ncols=3, xmin=0, xmax=10, ymin=0, ymax=10)
values(r) <- 1:ncell(r)
s <- rast(nrows=25, ncols=30, xmin=1, xmax=11, ymin=-1, ymax=11)
x <- resample(r, s, method="bilinear")
plot(r, col = pal_white_blue,main = "Original raster")
plot(x, col = pal_white_blue,main = "Resampled raster")

```


--- 

## Raster - vector interactions 

```{r}
CZProfile <- RCzechia::vyskopis(format="actual",cropped = F)
CZProfile
# get Czechia state boundaries
CZbound <- RCzechia::republika(resolution = "low")
#
# plot raster
plot(CZProfile)
plot(CZbound,
     border = "red",
     col = NA,
     add = TRUE)
#
# crop raster image (to a bounding box of CZ borders)
cropped <- terra::crop(CZProfile, CZbound) # crop raster to the bbox of polygon
#
# plot cropped raster
plot(cropped)
plot(CZbound,
     border = "red",
     col = NA,
     add = TRUE)
#
# mask to a defined polygon boundary (drop rest)
masked <- terra::mask(cropped, CZbound)
#
# plot masked raster
plot(masked)
#
inv_mask <- terra::mask(cropped, CZbound, inverse = TRUE)
#
# plot masked raster
plot(inv_mask)

```


## Raster data extraction

```{r}
set.seed(2345)
CZspots <- spatSample(
  masked,
  size = 5,
  method = "random",
  as.points = TRUE,
  na.rm = TRUE,
  values = FALSE
)
plot(masked,main = "CZ elevation")
points(CZspots,pch = 21,bg = "red",col = "black",cex = 1.5)
elevation <- terra::extract(masked, CZspots)
```

## D5 elevation plot.


```{r}

sample_every_m <- 250
simplify_tolerance_m <- 300

# If D5 contains many very short pieces such as ramps/crossings,
# increase this, e.g. 500 or 1000.
min_part_length_m <- 500

# ------------------------------------------------------------
# 1. Define bounding box
# ------------------------------------------------------------

# Larger region around western / central Czechia
# You can also use getbb("Czech Republic"), but a smaller bbox is faster.
bbox_d5 <- c(
  xmin = 12.3,
  ymin = 49.4,
  xmax = 15.0,
  ymax = 50.3
)

# ------------------------------------------------------------
# 2. Query OSM for D5 motorway
# ------------------------------------------------------------

try(osm_d5 <- opq(bbox = bbox_d5) |>
  add_osm_feature(key = "highway", value = "motorway") |>
  add_osm_feature(key = "ref", value = "D5") |>
  osmdata_sf()
)

# ------------------------------------------------------------
# 3. Extract line geometries
# ------------------------------------------------------------
try(D5_osm <- osm_d5$osm_lines)
# ------------------------------------------------------------
# If osm database is not responding (heavy traffic is common)
# you can read the D5_osm next 
# (the OSM dowload has 4MB+, so just the linestring are stored)
D5_osm <- st_read("data/D5_osm.gpkg")
# ------------------------------------------------------------


# Quick check
names(D5_osm)
nrow(D5_osm)
st_geometry_type(D5_osm)


D5_osm_main <- D5_osm |>
  st_transform(5514) |>
  st_make_valid() |>
  summarise() |>
  st_line_merge()


D5_geom <- D5_osm_main |>
  st_geometry() |>
  st_union() |>
  st_collection_extract("LINESTRING") |>
  st_line_merge() 

D5 <- st_sf(
  geometry = st_cast(D5_geom, "LINESTRING")
) |>
  mutate(
    part_id = row_number(),
    part_len_m = as.numeric(st_length(geometry))
  ) |>
  filter(part_len_m >= min_part_length_m)
# 
D5 <- D5[2,] # direction: Border >> Prague
D5wgs84 <- D5 |> 
  st_transform(4326)

plot(masked,main = "CZ elevation")
lines(D5wgs84,col = "red",lwd = 2)


# ------------------------------------------------------------
# Sample points every 250 m along the smoothed D5 line
# ------------------------------------------------------------

D5_length_m <- as.numeric(st_length(D5))

dist_m <- seq(0, D5_length_m, by = sample_every_m)

# Include final endpoint
if (max(dist_m) < D5_length_m) {
  dist_m <- c(dist_m, D5_length_m)
}

D5_points_geom <- st_line_sample(
  D5,
  sample = dist_m / D5_length_m
)

D5_points <- st_sf(
  id = seq_along(dist_m),
  dist = dist_m,
  dist_km = dist_m / 1000,
  geometry = st_cast(D5_points_geom, "POINT")
)


# ------------------------------------------------------------
# 6. Extract elevation from raster object masked
# ------------------------------------------------------------

# Transform points to raster CRS
D5_points_raster_crs <- st_transform(D5_points, crs(masked))

# Extract raster values
D5_elev <- terra::extract(
  masked,
  terra::vect(D5_points_raster_crs)
)

# Use the first raster value column returned by terra::extract()
elev_col <- setdiff(names(D5_elev), "ID")[1]

D5_points$elevation <- D5_elev[[elev_col]]


# ------------------------------------------------------------
# 7. Plot map check
# .. you can see that the simplification is far from perfect,
#    however, this should not have major consequences for the
#    elevation plot.
# ------------------------------------------------------------

ggplot() +
  geom_sf(data = D5, colour = "red", linewidth = 1) +
  geom_sf(data = D5_points, colour = "black", size = 0.4) +
  labs(
    title = "D5 highways"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5)
  )


# ------------------------------------------------------------
# 8. Plot elevation profile
# ------------------------------------------------------------

D5_points |>
  arrange(dist_km) |>
  ggplot(aes(x = dist_km, y = elevation)) +
  geom_line(linewidth = 0.8, colour = "steelblue") +
  labs(
    x = "Distance along D5, from state border (left) to Prague (right), in km",
    y = "Elevation",
    title = "Elevation profile of the D5 highway in Czechia"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```


## elevation to H3 (Uber) hexagons, level 6


```{r}
# prepare H3 hexagonss
CZ_buff <- CZbound %>% 
  st_transform(crs = 3587) %>% 
  st_buffer(dist = 800) %>% 
  st_transform(crs = 4326)
  
hex_ids <- h3jsr::polygon_to_cells(CZ_buff, res = 6)
hex_sf <- h3jsr::cell_to_polygon(hex_ids, simple = F)
hex_sf <- st_transform(hex_sf, st_crs(CZbound))
hex_sf <- hex_sf[st_intersects(hex_sf, CZbound, sparse = F), ]
hex_sf <- st_intersection(hex_sf, CZbound)
plot(hex_sf[,1])
#
hex_sf$elevation <- exactextractr::exact_extract( # faster than terra::extract()
  x = masked, # source, masked raster with CZ profile
  y = hex_sf, # target, sf polygons 
  fun = "mean", # calculation of average built-up - can be used for air pollution, etc.
  weights = "area",
  coverage_area = TRUE,
  progress = FALSE # hide console progress visualization
) 

## elevation to H3 grid
ggplot(data = hex_sf, aes(fill = elevation, label = round(elevation))) +
  geom_sf(lwd = 1/3) +
  scale_fill_viridis_c(
    name = "Terrain \nprofile \nH3 lev.6 \n(2321 h.)"
  ) +
  theme(
    axis.title = element_blank(),
    plot.caption = element_text(face = "italic"))
```



## Rasterization and vectorization


```{r}
# Rasterization
cycleHire <- spData::cycle_hire_osm
mapview::mapview(cycleHire)
cycleHire_projected <- st_transform(cycleHire, "EPSG:27700")
raster_template <- rast(ext(cycleHire_projected), resolution = 1000,
                       crs = crs(cycleHire_projected))
# presence/absence
ch_raster1 <- rasterize(cycleHire_projected, raster_template)
# different cycle hire locations have different numbers of bicycles 
# described by the capacity variable
# -> calculate total capacity in each grid cell 
ch_raster2 <- rasterize(cycleHire_projected, raster_template, 
                       field = "capacity", fun = sum, na.rm = TRUE)
#
ch_raster1a <- project(ch_raster1, "epsg:4326")
ch_raster2a <- project(ch_raster2, "epsg:4326")
#
mapview::mapview(ch_raster1a,legend=FALSE)
mapview::mapview(ch_raster2a,col.regions = viridisLite::viridis)


# Vectorization
CZcontour <- as.contour(masked)
CZcontour
CZcontour_sf <- st_as_sf(CZcontour)
CZcontour_sf
# Contour plot
plot(masked)
lines(CZcontour)
# Other version of the plot
br <- c(0, 300, 600,900,1200,1604)
plot(masked, breaks=br)
lines(as.contour(masked, levels=br))

```