---
title: "Block 1 R code"
format: html
---

## Block 1 - presentation: `R` code

```{r}
#| warning: false
#| message: false
library(RCzechia)
library(terra)
library(eurostat)
library(ggplot2)
library(RColorBrewer)
library(lwgeom)
library(tidyverse)
library(sf)
library(s2)
library(mapview)
library(rnaturalearth)
```

------------------------------------------------------------------------

## Three types of spatial data (rasters, vectors/polygons, spatial graphs)

```{r}
Elevation <- RCzechia::vyskopis(format="actual")
CZbound <- RCzechia::republika(resolution = "low")
Elevation <- terra::mask(Elevation,CZbound)
plot(Elevation)

CZcounties <- RCzechia::okresy(resolution = "low")
plot(CZcounties[,1])

CZrailways <- RCzechia::zeleznice()
CZrailways$ELEKTRIFIKACE <-as.factor(CZrailways$ELEKTRIFIKACE)
plot(CZrailways["ELEKTRIFIKACE"],lwd=2,reset=F)
plot(st_geometry(CZbound),add=T)

```


------------------------------------------------------------------------

## Unemployment in Central Europe: panel data

```{r}

ce_countries <- c("AT", "CZ", "DE", "HU", "PL", "SK")
years_keep   <- c(2016, 2019)

# NUTS2 geometries, GISCO 2010, 1:60M scale
nuts2 <- eurostat::get_eurostat_geospatial(
  nuts_level   = 2,
  year         = 2016,
  resolution   = "60",
  output_class = "sf",
  crs          = "4326"
) |>
  st_make_valid() |>
  dplyr::filter(CNTR_CODE %in% ce_countries)
# Country boundaries, NUTS0
nuts0 <- eurostat::get_eurostat_geospatial(
  nuts_level   = 0,
  year         = 2016,
  resolution   = "60",
  output_class = "sf",
  crs          = "4326"
) |>
  st_make_valid() |>
  dplyr::filter(CNTR_CODE %in% ce_countries)


# Eurostat unemployment rate
# Dataset: lfst_r_lfu3rt
# sex = total
# age = 20-64
# time = 2005 and 2012
# ------------------------------------------------------------
unemp <- eurostat::get_eurostat(
  id          = "lfst_r_lfu3rt",
  time_format = "num",
  type        = "code")

unemp <- unemp %>% 
  dplyr::filter(sex  == "T", age  == "Y20-64", isced11 == "TOTAL",
                TIME_PERIOD %in% years_keep)



# Join unemployment data to NUTS2 polygons
map_data <- nuts2 |>
  left_join(
    unemp,
    by = c("NUTS_ID" = "geo")
  ) |>
  mutate(
    # Small enclosed city regions should be drawn last.
    # This replaces the old TopL logic for Berlin, Bremen, Prague, Vienna.
    top_layer = if_else(NUTS_ID %in% c("DE30", "DE50", "CZ01", "AT13"), 1L, 0L)
  ) |>
  arrange(top_layer)


# Horizontal two-panel plot
p_horizontal <- ggplot() +
  geom_sf(
    data = map_data,
    aes(fill = values),
    color = "grey50",
    linewidth = 0.1
  ) +
  geom_sf(
    data = nuts0,
    fill = NA,
    color = "black",
    linewidth = 0.4
  ) +
  scale_fill_gradientn(
    name = "% rate of\nunemp.",
    colours = brewer.pal(9, "Reds"),
    na.value = "white"
  ) +
  # coord_sf(crs = st_crs(4326)) +
  facet_wrap(~ TIME_PERIOD, nrow = 1) +
  labs(
    title = "Unemployment rate, years 2016 and 2019"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    panel.grid.major = element_line(color = "grey90", linewidth = 0.2),
    # strip.text = element_text(face = "bold"),
    legend.position = "right",
    panel.spacing.x = unit(1, "cm")
  )

p_horizontal
```




------------------------------------------------------------------------

## Creating a point geometry in `R`

- approximate location of room NB432

```{r}
p <- sf::st_as_sfc("POINT(14.442408 50.083300)", crs = 4326)
p[[1]]
p
mapview::mapview(p)
```

------------------------------------------------------------------------

## Polygon projection (Polar region)

```{r}
ne <- countries110 %>% 
  st_as_sf() %>%
  st_geometry() %>% 
  st_transform("+proj=eqc")
plot(ne, axes = TRUE)

# sfc object: polygon is created
# note how the first and last elements are identical 
pol = st_as_sfc("POLYGON((-150 -65, 0 -62, 120 -78, -150 -65))", crs = 4326)
pol
pol <- st_transform(pol,"+proj=eqc")
# sfc object: point (South pole) is created
pole = st_as_sfc("POINT(0 -90)", crs = 4326) %>% 
  st_transform("+proj=eqc")

# does the polygon contain South pole?
st_intersects(pole,pol)


plot(ne, axes = TRUE, reset=F)
plot(pol, add=T, col = "red")


world_antarctic <- countries110 |>
  dplyr::filter(ADMIN == "Antarctica") |>
  st_geometry() |>
  st_transform("+proj=eqc")  

# Orthographic view centered on South Pole
ortho <- "+proj=ortho +lat_0=-90 +lon_0=0"

plot(
  st_transform(world_antarctic, ortho),
  xlim = c(-3e6, 3e6),
  ylim = c(-3e6, 3e6),
  col = "white",
  border = "black"
)

plot(
  st_transform(
    st_segmentize(
      pol,
      units::set_units(20, km)
    ),
    ortho
  ),
  add = TRUE,
  col = "red",
  border = "black"
)

```

Another way to interpret & visualize the polygon: instead of re-projecting the original shape (focusing on the border-lines of the original object) to polar coordinates, we can simply take the three vertices (of a triangle) and connect them with a *new* shape (that actually covers the South Pole). 

```{r}
pol2 = st_as_sfc("POLYGON((-150 -65, 0 -62, 120 -78, -150 -65))", crs = 4326)


plot(
  st_transform(world_antarctic, ortho),
  xlim = c(-3e6, 3e6),
  ylim = c(-3e6, 3e6),
  col = "white",
  border = "black"
)

plot(
  st_transform(
    st_segmentize(
      pol,
      units::set_units(20, km)
    ),
    ortho
  ),
  add = TRUE,
  col = "red",
  border = "black"
)

plot(
  st_transform(pol2,ortho),
  add = TRUE,
  col = "darkgreen",
  border = "black"
)

plot(
  st_transform(pole,ortho),
  add = TRUE,
  col = "yellow",
  border = "yellow"
)
```
------------------------------------------------------------------------
## Germany in three different projections

```{r}
allNUTSmap <- eurostat::get_eurostat_geospatial(year="2013")
# Example of map for Germany - NUTS 2, WGS84
Germany_NUTS2 <- allNUTSmap %>%   
  dplyr::filter(LEVL_CODE == 2 & CNTR_CODE %in% ("DE")) 
ggplot() + 
  geom_sf(data = Germany_NUTS2) +
  theme_bw()
#
# LAEA geometry
laea = st_crs("+proj=laea +lat_0=51 +lon_0=11") 
GermanyLAEA <- st_transform(Germany_NUTS2,laea)
ggplot() + 
  geom_sf(data = GermanyLAEA)+
  theme_bw()
#
# LAEA geometry optimized for Iceland
GermanyLAEAIceland <- st_transform(Germany_NUTS2,crs=9947)
ggplot() + 
  geom_sf(data = GermanyLAEAIceland)+
  theme_bw()
```

------------------------------------------------------------------------

## Unemployment Germany: choropleth/infomap

```{r}

# example continued from CE panel data (map_data)

# Join unemployment data to NUTS2 polygons
DE <- map_data %>% 
  dplyr::filter(CNTR_CODE == "DE", TIME_PERIOD == 2016) %>% 
  dplyr::select(geo,values) # geometry is not lost, DE has 3 columns


# Horizontal two-panel plot
ggplot(DE) +
  geom_sf(aes(fill = values))+
  scale_fill_gradient(
    name = "% rate of\nunemp.",
    low = "white",
    high = "red")+
  theme_bw()
```


------------------------------------------------------------------------

## Seven basic vector-based map elements

```{r}
# 1 point
p <- st_point(0:1)
plot(p, pch = 16)
title("point")
box(col = 'grey')

# 2 multipoint
mp <- st_multipoint(rbind(c(1,1), c(2, 2), c(4, 1), c(2, 3), c(1,4)))
plot(mp, pch = 16)
title("multipoint")
box(col = 'grey')

# 3 linestring
ls <- st_linestring(rbind(c(1,1), c(5,5), c(5, 6), c(4, 6), c(3, 4), c(2, 3)))
plot(ls, lwd = 2)
title("linestring")
box(col = 'grey')

# 4 multilinestring
mls <- st_multilinestring(list(
  rbind(c(1,1), c(5,5), c(5, 6), c(4, 6), c(3, 4), c(2, 3)),
  rbind(c(3,0), c(4,1), c(2,1))))
plot(mls, lwd = 2)
title("multilinestring")
box(col = 'grey')

# 5 polygon
po <- st_polygon(list(rbind(c(2,1), c(3,1), c(5,2), c(6,3), c(5,3), c(4,4), c(3,4), c(1,3), c(2,1)),
    rbind(c(2,2), c(3,3), c(4,3), c(4,2), c(2,2))))
plot(po, border = 'black', col = '#ff8888', lwd = 2)
title("polygon")
box(col = 'grey')

# 6 multipolygon
mpo <- st_multipolygon(list(
    list(rbind(c(2,1), c(3,1), c(5,2), c(6,3), c(5,3), c(4,4), c(3,4), c(1,3), c(2,1)),
        rbind(c(2,2), c(3,3), c(4,3), c(4,2), c(2,2))),
    list(rbind(c(3,7), c(4,7), c(5,8), c(3,9), c(2,8), c(3,7)))))
plot(mpo, border = 'black', col = '#ff8888', lwd = 2)
title("multipolygon")
box(col = 'grey')

# 7 geometrycollection
gc <- st_geometrycollection(list(po, ls + c(0,5), st_point(c(2,5)), st_point(c(5,4))))
plot(gc, border = 'black', col = '#ff6666', pch = 16, lwd = 2)
title("geometrycollection")
box(col = 'grey')

```

------------------------------------------------------------------------
